Modeling Two-Dimensional Vesicle Hydrodynamics with Hydrophobic Attraction Potential Simulations

### 2021-08-06 doc/main_jfm.tex
worked on Secs 2 and 3

\section{Governing  Equations}
\subsection{\label{mobility}Mobility Problem}
We consider a suspension of $N_b$ particles suspended in a
two-dimensional unbounded domain. The boundary of each particle is
denoted by $\Gamma_i$ so that $\bd \Omega = \Gamma_1 \cup \Gamma_2 \cup
\cdots \cup \Gamma_{N_b}$. Assuming the flow is Stokesian, the governing
equations are
\begin{alignat}{3}
  -\mu \Delta \uu + \nabla p &= \mathbf{0}, \qquad \qquad 
    && \xx \in \Omega, \\
  \nabla\cdot \uu &= 0, \qquad && \xx \in \Omega, \\
  \uu - \uu_\infty &\to \mathbf{0}, && |\xx| \to \infty,
\end{alignat}
%
where $\uu$ is the velocity, $\mu$ is viscosity, $p$ is the pressure,
and $\uu_\infty$ is the background flow. Since each particle $\Gamma_i$
with center $\aa_i$ is a rigid body, its velocity satisfies 
\begin{align}
  \vv(\xx) = \vv_i + \omega_i (\xx - \aa_i)^\perp, \quad 
    \xx \in \Gamma_i,
\end{align}
where $\vv_i$ is its translational velocity and $\omega_i$ is its
angular velocity.  Here, $\langle x, y \rangle^{\perp} = \langle -y, x
\rangle$. Therefore, the no-slip boundary condition on each particle is
\begin{align}
  \uu(\xx) = \vv_i + \omega_i (\xx - \aa_i)^\perp, \quad
    \xx \in \Gamma_i.
\end{align}
To determine the translational and angular velocities of each particle,
we assume some imposed forces $\FF_i$ and torques $T_i$ acting on each
particle. Since the small particles are intertialess, force balance
gives 
\begin{alignat}{2}
  \label{eqn:force}
  \FF_i &+ \int_{\Gamma_i} \ssigma \cdot \nnu_i \, ds = \mathbf{0},
  && i=1,\ldots,N_b,\\
  \label{eqn:torque}
  T_i &+ \int_{\Gamma_i} (\xx - \aa_i)^\perp \cdot 
    (\ssigma \cdot \nnu_i) \, ds = 0 \qquad && i=1,\ldots,N_b,
\end{alignat}
where $\ssigma = -p \mathbf{I} + \mu \left(\nabla \uu + \nabla \uu^T
\right)$ is the hydrodynamic stress tensor (pressure tensor).

\subsection{Imposed Forces}
The imposed forces and torques contain two parts. 
The first part comes from the hydrophobic attraction potential (HAP).
This potential was introduced in \cite{Fu20} and
models hydrophobic attraction by solving the screened Laplace equation boundary value problem
\begin{alignat}{2}
  \label{SL}
-\rho^2 \Delta u + u &=0,            && \xx \in \Omega,\\
\label{SLbc}
u(\xx) &= f_i(\xx),\qquad  && \xx \in \Gamma_i, \\
\label{SLff}
u &\to 0,                          &&|\xx| \to \infty,
\end{alignat}
where $f_i$ is material label with $f_i = 0$ ($1$) representing
hydrophilic (hydrophobic) portions of the surface, and $\rho$ is the
decay length of attraction.  The forces and torques for this attraction
are 
\begin{align}
  \label{eqn:hydrophobicAttraction}
  \FF_i^{\text{hydro}} = \int_{\Gamma_i} {\bf T}\cdot \nnu_i \, dS, 
    \quad 
  T_i^{\text{hydro}} = \int_{\Gamma_i} (\xx - \aa_i)^{\perp} \cdot ({\bf T} \cdot \nnu_i) dS,
\end{align}
where
\begin{align}
  \label{eq:stress}
\mathbf{T}
= \gamma\rho^{-1}u^2 \mathbf{I} + 2\rho\gamma \left(\tfrac{1}{2}|\nabla
  u|^2 \mathbf{I} - \nabla u\otimes \nabla u\right)
\end{align}
is the hydrophobic stress tensor. The parameter $\gamma$ is interfacial
tension. The forces $\FF_i^{\text{hydro}}$ and torques
$T_i^{\text{hydro}}$ are responsible for forming particle aggregates
that partially sequester the hydrophobic portions of the particle
surface.

The second part of the imposed forces and torques comes from repulsion
between proximal particles. We define 
\begin{align}
%  \label{eqn:REPUL}
%  \FF_i^{\text{repul}} = \sum_{i=1}^{N_b} \frac{\xx_{ij} -
%  \yy_{ij}}{\rho_{ij}} P(\rho_{ij}),\quad 
%  T_i^{\text{repul}} = \sum_{i=1}^{N_b} (\xx_{ij} - \aa_i)^{\perp}\cdot
%  \frac{\xx_{ij} - \yy_{ij}}{\rho_{ij}} P(\rho_{ij})
%  \\
%
  \label{eqn:REPULforce}
  \FF_i^{\text{repul}} &= \sum_{j \neq i} \int_{\Gamma_j}
    \frac{\xx - \yy}{|\xx - \yy|} P(|\xx - \yy|)\, dS, \\
  \label{eqn:REPULtorque}
  T_i^{\text{repul}} &= \sum_{j \neq i} \int_{\Gamma_j}
    (\xx - \aa_i)^{\perp} \cdot 
    \frac{\xx - \yy}{|\xx - \yy|} P(|\xx - \yy|)\, dS, 
\end{align}
\todo[inline]{Repulsion force is a function of only the distance between
the Janus particles instead of the integral around the particles.}
where $\xx \in \Gamma_i$ and $|\xx - \yy|$ is the distance between the
pair of particles. The repulsion profile $P(r)$ is zero for distances
$r$ larger than a repulsion length scale $\rho_0$. As
such,~\eqref{eqn:REPULforce} and~\eqref{eqn:REPULtorque} ignore
particles $\Gamma_j$ outside a tubular neighborhood of $\Gamma_i$. For
$0 \leq r < \rho_0,$ we use $P(r) = M(1 - \sin(r/\rho_0))$ where $M$ is
sufficiently large to prevent particle collisions. The total imposed
force and torque are
\begin{align}
  \FF_i = \FF_i^{\text{hydro}} + \FF_i^{\text{repul}},\quad
  T_i = T_i^{\text{hydro}} + T_i^{\text{repul}}, \qquad
  i=1,\ldots,N_b.
\end{align}

\subsection{Time Marching and Nondimensionalizations}
By solving the mobility problem, we obtain translational and angular
velocities of the $N_b$-body system. A second-order Adam-Bashforth
scheme updates the particle positions and orientations. By including the
repulsion force~\eqref{eqn:REPULforce} and
torque~\eqref{eqn:REPULtorque}, particle collisions are avoided even
when using a relatively large time step.

@@

%%%
\section{Integral Equation Method}
\label{sec:IEM}
Since the fluid velocity is assumed to be Stokesian, it can be written
in terms of layer potentials. Following previous work of~\cite{Lukas19},
we represent the velocity as the sum of a double-layer potential and the
$N_b$ Stokeslets and Rotlets
\begin{align}
  \label{eqn:velocity}
  \uu(\xx) = \uu_\infty(\xx) + \DD[\eeta](\xx) + 
    \sum_{i=1}^{N_b} \left(\SS(\xx,\aa_i) \FF_i + 
    \RR(\xx,\aa_i) T_i\right), \quad \xx \in \Omega.
\end{align}
The double-layer potential is
\begin{align}
  \DD[\eeta](\xx) = \frac{1}{\pi} \int_{\bd\Omega} 
    \frac{\rr \cdot \nnu}{\rho^2} \frac{\rr \otimes \rr}{\rho^2}
    \eeta(\yy) \, ds_\yy,
\end{align}
where $\rr = \xx - \yy$ and $\rho = |\rr|$. The Stokeslet and Rotlet
centered at $\aa_i$ are
\begin{align}
  \SS(\xx,\aa_i)\FF_i &= \frac{1}{4\pi} \left(-\log(\rho) + 
    \frac{\rr \otimes \rr}{\rho^2} \FF_i \right), \\
  \RR(\xx,\aa_i)T_i &= \frac{1}{4\pi} \frac{\rr^\perp}{\rho^2} T_i,
\end{align}
where $\rr = \xx - \aa_i$ and $\rho = |\rr|$. The Stokeslet is
torque-free and has force $\FF$, and the Rotlet force-free and has
torque $T$. Therefore, the velocity~\eqref{eqn:velocity} satisfies the
total force~\eqref{eqn:force} and torque~\eqref{eqn:torque} conditions
if the double-layer potential $\DD[\eeta]$ is force- and torque-free.
Matching the limit of~\eqref{eqn:velocity} with the rigid body motion,
and imposing that $\DD[\eeta]$ is force- and torque-free, the density
function $\eeta$, Stokeslet strength, $\FF_i$, and Rotlet strength $T_i$
satisfy
\begin{alignat}{3}
  \nonumber
  \vv_j + \omega_j (\xx - \aa_j)^\perp &= \uu_\infty(\xx)
    -\frac{1}{2} \eeta(\xx) + \DD[\eeta](\xx) \\
  \label{eqn:SKIE}
    &+ \sum_{i=1}^{N_b} 
    \left(\SS(\xx,\aa_i) \FF_i + \RR(\xx,\aa_i) T_i\right),
    \quad &&\xx \in \bd \Gamma_j, \\
  \label{eqn:mobility1}
  \int_{\bd\Gamma_i} \eeta \cdot \nnu_i \, ds &= {\bf 0}, 
  &&i = 1,\ldots,N_b, \\
  \label{eqn:mobility2}
  \int_{\bd\Gamma_i} \eeta\times(\xx-\aa_i)^\perp \cdot \nnu ds &= 0,
  &&i = 1,\ldots,N_b.
\end{alignat}
%
%The mobility problem for $N_b$-body system in a Stokes flow leads to a second-kind boundary integral equations (BIE) with a density function $\eta({\bf x}$ that is given by \cite{Lukas19}
%\begin{equation}
%\label{mobility}
%\begin{aligned}
%{\bf v}_i + &\omega_i\times({\bf x}-{\bf a}_i)\\
% =& -\frac12\eta({\bf x}) + D[\eta]({\bf x}) \\
%&+ \sum_{i=1}^N\big({\bf S}({\bf x}, {\bf a}_i){\bf F}_i+ {\bf R}({\bf x}, {\bf a}_i)t_i\big)+{\bf u}_\infty, \quad {\bf x}\in\Gamma \\
%\end{aligned}
%\end{equation}
%%
%with two constraints 
%%
%\begin{center}
%\begin{equation}
%\label{mobility2}
%\int_{\partial \Gamma_i}\eta\cdot \nnu ds = {\bf 0},
%\end{equation}
%\end{center}
%%
%\begin{center}
%\begin{equation}
%\label{mobility3}
%\int_{\partial \Gamma_i}\eta\times({\bf x}-{\bf a}_i)\cdot \nnu ds ={\bf 0},
%\end{equation}
%\end{center}
%%
%where the two-dimensional Stokeslet and the Rotlet are
%%
%\begin{gather}
%{\bf S}({\bf x}, {\bf a}) = \frac{1}{4\pi}\left(-\log(|({\bf x}-{\bf a})|){\bf I}+\frac{({\bf x}-{\bf a})\otimes({\bf x}-{\bf a})}{|({\bf x}-{\bf a})|^2}\right) \\ 
%{\bf R}({\bf x},{\bf a})=\frac{({\bf x}-{\bf a})^\perp}{4\pi|({\bf x}-{\bf a})|^2}.
%\end{gather}
%
%Here ${\bf r}={\bf x}-{\bf a}$ and $\rho=|{\bf r}|$.
Given a collection of particles with forces and torques calculated from
the hydrophobic interactions and the short range repulsion,
equations~\eqref{eqn:SKIE}--\eqref{eqn:mobility2} form a solvable set of
integral equations for the translational velocities $\vv_i$, the angular
velocities $\omega_i$, and the density function $\eeta$.

\todo[inline]{Talk about how to solve numerically. Quadrature, time
stepping, collocation, etc.}
This system of equations is then solved iteratively with block-diagonal
preconditioned GMRES.

>>>>

\section{Governing  Equations}
\subsection{\label{mobility}Mobility Problem}
We consider a suspension of $N_b$ particles suspended in a
two-dimensional unbounded domain. The boundary of each particle is
denoted by $\Gamma_i$ so that $\bd \Omega = \Gamma_1 \cup \Gamma_2 \cup
\cdots \cup \Gamma_{N_b}$. Assuming the flow is Stokesian, the governing
equations are
\begin{alignat}{3}
  -\mu \Delta \uu + \nabla p &= \mathbf{0}, 
    && \xx \in \Omega, \\
  \nabla\cdot \uu &= 0, \qquad && \xx \in \Omega, \\
  \uu - \uu_\infty &\to \mathbf{0}, && |\xx| \to \infty,
\end{alignat}
%
where $\uu$ is the velocity, $\mu$ is viscosity, $p$ is the pressure,
and $\uu_\infty$ is the background flow. Since each particle $\Gamma_i$
with center $\aa_i$ is a rigid body, its velocity satisfies 
\begin{align}
  \vv(\xx) = \vv_i + \omega_i (\xx - \aa_i)^\perp, \quad 
    \xx \in \Gamma_i,
\end{align}
where $\vv_i$ is its translational velocity and $\omega_i$ is its
angular velocity. Here, $\langle x, y \rangle^{\perp} = \langle -y, x
\rangle$. Therefore, the no-slip boundary condition on each particle is
\begin{align}
  \uu(\xx) = \vv_i + \omega_i (\xx - \aa_i)^\perp, \quad
    \xx \in \Gamma_i.
\end{align}
To determine the translational and angular velocities of each particle,
we assume some imposed forces $\FF_i$ and torques $T_i$ acting on each
particle. Since the small particles are intertialess, force balance
gives 
\begin{alignat}{2}
  \label{eqn:force}
  \FF_i &+ \int_{\Gamma_i} \ssigma \cdot \nnu_i \, ds = \mathbf{0},
  && i=1,\ldots,N_b,\\
  \label{eqn:torque}
  T_i &+ \int_{\Gamma_i} (\xx - \aa_i)^\perp \cdot 
    (\ssigma \cdot \nnu_i) \, ds = 0, \qquad && i=1,\ldots,N_b,
\end{alignat}
where $\ssigma = -p \mathbf{I} + \mu \left(\nabla \uu + \nabla \uu^T
\right)$ is the hydrodynamic stress tensor (pressure tensor). The
process of finding the translational and angular velocities given a
force and torque is referred to as the mobility problem.

\subsection{Imposed Forces}
The imposed forces and torques contain two parts. The first part comes
from the hydrophobic attraction potential (HAP). This potential was
introduced in~\cite{Fu20} and models hydrophobic attraction by solving
the screened Laplace equation boundary value problem
\begin{alignat}{2}
  \label{SL}
-\rho^2 \Delta u + u &=0,            && \xx \in \Omega,\\
\label{SLbc}
u(\xx) &= f_i(\xx),\qquad  && \xx \in \Gamma_i, \\
\label{SLff}
u &\to 0,                          &&|\xx| \to \infty,
\end{alignat}
where $f_i$ is a material label with $f_i = 0$ ($1$) representing
hydrophilic (hydrophobic) portions of the surface, and $\rho$ is the
decay length of attraction. The forces and torques for this attraction
are 
\begin{align}
  \label{eqn:hydrophobicAttraction}
  \FF_i^{\text{hydro}} = \int_{\Gamma_i} {\bf T}\cdot \nnu_i \, dS, 
    \quad 
  T_i^{\text{hydro}} = \int_{\Gamma_i} (\xx - \aa_i)^{\perp} \cdot ({\bf T} \cdot \nnu_i) dS,
\end{align}
where
\begin{align}
  \label{eq:stress}
\mathbf{T}
= \gamma\rho^{-1}u^2 \mathbf{I} + 2\rho\gamma \left(\tfrac{1}{2}|\nabla
  u|^2 \mathbf{I} - \nabla u\otimes \nabla u\right)
\end{align}
is the hydrophobic stress tensor. The parameter $\gamma$ is interfacial
tension. The forces $\FF_i^{\text{hydro}}$ and torques
$T_i^{\text{hydro}}$ are responsible for forming particle aggregates
that partially sequester the hydrophobic portions of the particle
surface.

The second part of the imposed forces and torques comes from repulsion
between proximal particles. Given a pair particles indexed with $i$ and
$j$, we find the two points $\tilde{\xx}_i \in \Gamma_i$ and
$\tilde{\xx}_j \in \Gamma_j$ that are closest to one another. We then
define the repulsion force and torque
\begin{align}
%  \label{eqn:REPUL}
%  \FF_i^{\text{repul}} = \sum_{i=1}^{N_b} \frac{\xx_{ij} -
%  \yy_{ij}}{\rho_{ij}} P(\rho_{ij}),\quad 
%  T_i^{\text{repul}} = \sum_{i=1}^{N_b} (\xx_{ij} - \aa_i)^{\perp}\cdot
%  \frac{\xx_{ij} - \yy_{ij}}{\rho_{ij}} P(\rho_{ij})
%  \\
%
  \label{eqn:REPULforce}
  \FF_i^{\text{repul}} &= \sum_{j \neq i} 
    \frac{\tilde{\xx}_i - \tilde{\xx}_j}
    {|\tilde{\xx}_i - \tilde{\xx}_j|} 
    P'(|\tilde{\xx}_i - \tilde{\xx}_j|), \\
  \label{eqn:REPULtorque}
  T_i^{\text{repul}} &= \sum_{j \neq i} 
    (\tilde{\xx}_i - \aa_i)^{\perp} \cdot 
    \frac{\tilde{\xx}_i - \tilde{\xx}_j}
    {|\tilde{\xx}_i - \tilde{\xx}_j|} 
    P'(|\tilde{\xx}_i - \tilde{\xx}_j|).
\end{align}
The repulsion profile $P(r)$ is zero for distances $r$ larger than a
repulsion length scale $\rho_0$. As such,~\eqref{eqn:REPULforce}
and~\eqref{eqn:REPULtorque} ignore particles $\Gamma_j$ outside a
tubular neighborhood of $\Gamma_i$. For $0 \leq r < \rho_0,$ we use
$P(r) = M(1 - \sin(r/\rho_0))$ where $M$ is sufficiently large to
prevent particle collisions. Then, the total imposed force and torque
are
\begin{align}
  \FF_i = \FF_i^{\text{hydro}} + \FF_i^{\text{repul}},\quad
  T_i = T_i^{\text{hydro}} + T_i^{\text{repul}}, \qquad
  i=1,\ldots,N_b.
\end{align}

\subsection{Time Marching}
By solving the mobility problem, we obtain translational and angular
velocities of the $N_b$-body system. A second-order Adam-Bashforth
scheme updates the particle positions and orientations. By including the
repulsion force~\eqref{eqn:REPULforce} and
torque~\eqref{eqn:REPULtorque}, particle collisions are avoided even
when using a relatively large time step.

@@

%%%
\section{Integral Equation Method}
\label{sec:IEM}
Since the fluid velocity is assumed to be Stokesian, it can be written
in terms of layer potentials. Following previous work of~\cite{Lukas19},
we represent the velocity as the sum of a double-layer potential and the
$N_b$ Stokeslets and Rotlets
\begin{align}
  \label{eqn:velocity}
  \uu(\xx) = \uu_\infty(\xx) + \DD[\eeta](\xx) + 
    \sum_{i=1}^{N_b} \left(\SS(\xx,\aa_i) \FF_i + 
    \RR(\xx,\aa_i) T_i\right), \quad \xx \in \Omega.
\end{align}
The double-layer potential is
\begin{align}
  \DD[\eeta](\xx) = \frac{1}{\pi} \int_{\bd\Omega} 
    \frac{\rr \cdot \nnu}{\rho^2} \frac{\rr \otimes \rr}{\rho^2}
    \eeta(\yy) \, ds_\yy,
\end{align}
where $\rr = \xx - \yy$ and $\rho = |\rr|$. The Stokeslet and Rotlet
centered at $\aa_i$ are
\begin{align}
  \SS(\xx,\aa_i)\FF_i &= \frac{1}{4\pi} \left(-\log(\rho) + 
    \frac{\rr \otimes \rr}{\rho^2} \FF_i \right), \\
  \RR(\xx,\aa_i)T_i &= \frac{1}{4\pi} \frac{\rr^\perp}{\rho^2} T_i,
\end{align}
where $\rr = \xx - \aa_i$ and $\rho = |\rr|$. The Stokeslet is
torque-free and has force $\FF_i$ while the Rotlet force-free and has
torque $T_i$. Therefore, the velocity~\eqref{eqn:velocity} satisfies the
total force~\eqref{eqn:force} and torque~\eqref{eqn:torque} conditions
if the double-layer potential $\DD[\eeta]$ is force- and torque-free.
Matching the limit of~\eqref{eqn:velocity} with the rigid body motion,
and imposing that $\DD[\eeta]$ is force- and torque-free, the density
function $\eeta$, translational velocity $\vv_i$, and angular velocity
$\omega_i$ satisfy
\begin{alignat}{3}
  \nonumber
  \vv_i + \omega_i (\xx - \aa_i)^\perp &= \uu_\infty(\xx)
    -\frac{1}{2} \eeta(\xx) + \DD[\eeta](\xx) \\
  \label{eqn:SKIE}
    + \sum_{j=1}^{N_b} &
    \left(\SS(\xx,\aa_j) \FF_j + \RR(\xx,\aa_j) T_j\right),
    \quad &&\xx \in \bd \Gamma_i,\: i=1,\ldots,N_b, \\
  \label{eqn:mobility1}
  \int_{\bd\Gamma_i} \eeta \cdot \nnu_i \, ds &= {\bf 0}, 
  &&i = 1,\ldots,N_b, \\
  \label{eqn:mobility2}
  \int_{\bd\Gamma_i} \eeta\times(\xx-\aa_i)^\perp \cdot \nnu ds &= 0,
  &&i = 1,\ldots,N_b.
\end{alignat}

%Given a collection of particles with forces and torques calculated from
%the hydrophobic interactions and the short range repulsion,
%equations~\eqref{eqn:SKIE}--\eqref{eqn:mobility2} form a solvable set of
%integral equations for the translational velocities $\vv_i$, the angular
%velocities $\omega_i$, and the density function $\eeta$.

We discretize each of the $\Gamma_i$, $i=1,\ldots,N_b$, at $N$
discretization points. The trapezoid rule is used to approximate each of
the integrals in equations~\eqref{eqn:mobility1}
and~\eqref{eqn:mobility2}, and the double-layer potential
equation~\eqref{eqn:SKIE}. When two bodies are close, the double-layer
potential becomes nearly-singular, and we replace the double-layer
potential with an interpolation-based nearly-singular quadrature
rule~\cite{qua-bir2014}. After discretizing and applying appropriate
quadrature rules, the result is a $(2NN_b + 3N_b) \times (2NN_b + 3N_b)$
linear system whose condition number is independent of $N$, and this
linear system is solved using matrix-free GMRES. To reduce the number of
GMRES iterations, a block-diagonal preconditioner is used.
